---
title: "ToothGrowth Data Analysis"
author: "Pier Lorenzo Paracchini"
output: html_document
---
```{r settingsAndDependencies, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
Sys.setlocale("LC_ALL", "C")
```
##Overview
The goal is to analyze the ToothGrowth data in the R datasets package, performing some exploratory analysis, and then use confidence intervals and/or hypothesis tests to compare tooth growth by `supp` and `dose`. 

Description of the data from the documentation _"The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid)."_

##Getting and Exploring the data
```{r dataStructure, collpase = TRUE}
require(datasets)
rawData <- ToothGrowth
str(rawData)
```
The rawdata includes `r dim(rawData)[1]` obs. of `r dim(rawData)[2]` features. Features are `len` the tooth length (`numeric`), `supp` the supplement type (`factor`) and finally `dose` the dose in milligrams (`numeric`),


The rawdata includes 6 possible groups by `supp` and `dose` (see the plot below). For each supplement `supp` ("OJ" and "VC"), there are three different groups based on the possible `dose`s - 0.5mg, 1mg, 2mg.

```{r ref.label="basicPlot", echo=FALSE}
```

And here some basic statics about the sample tooth length for each group.
```{r ref.label="basicStatistics", echo=FALSE}
```
Supp | Dose | No of Sample | Sample mean | Sample SD 
---- | ---- | ---- | ---- | ----
"OJ" | 0.5 | `r n_OJ_05` | `r round(mean_OJ_05, 3)` | `r round(sd_OJ_05,3)` 
"OJ" | 1.0 | `r n_OJ_10` | `r round(mean_OJ_10, 3)` | `r round(sd_OJ_10,3)`
"OJ" | 2.0 | `r n_OJ_20` | `r round(mean_OJ_20, 3)` | `r round(sd_OJ_20,3)` 
"VC" | 0.5 | `r n_VC_05` | `r round(mean_VC_05, 3)` | `r round(sd_VC_05,3)` 
"VC" | 1.0 | `r n_VC_10` | `r round(mean_VC_10, 3)` | `r round(sd_VC_10,3)`
"VC" | 2.0 | `r n_VC_20` | `r round(mean_VC_20, 3)` | `r round(sd_VC_20,3)` 

##Confidence Intervals & Hypothesis Testing
###Basic Assumptions
Having each group has a limited number of sample (10 sample) then `t distribution` and `t confidence intervals` are going to be used. Being the guinea pigs used for each measurement distinct from each other within a group and between the groups then the __groups can be considered independent from each other__. Finally looking at the sample standard deviation for the different groups the assumption can be made to treat the diffent variances as equals. 
###Independent group t confident intervals
```{r ref.label="confidenceIntervals", echo=FALSE}
```

##Appendix
Code used to create the basic plot to visualize basic information of the rawData. 
```{r basicPlot}
require(graphics)
par(ps = 8, cex.lab = 0.8)
coplot(len ~ dose | supp, data = rawData, panel = panel.smooth, xlab = "ToothGrowth data: length vs dose, given type of supplement")
```

Code used to calculate basic statistics
```{r basicStatistics}
##OJ - Orange Juice
s_lenData_OJ_05 <- rawData[rawData$supp == "OJ" & rawData$dose == 0.5,]$len
n_OJ_05 <- length(s_lenData_OJ_05)
mean_OJ_05 <- mean(s_lenData_OJ_05)
sd_OJ_05 <- sd(s_lenData_OJ_05)

s_lenData_OJ_10 <- rawData[rawData$supp == "OJ" & rawData$dose == 1.0,]$len
n_OJ_10 <- length(s_lenData_OJ_10)
mean_OJ_10 <- mean(s_lenData_OJ_10)
sd_OJ_10 <- sd(s_lenData_OJ_10)

s_lenData_OJ_20 <- rawData[rawData$supp == "OJ" & rawData$dose == 2.0,]$len
n_OJ_20 <- length(s_lenData_OJ_20)
mean_OJ_20 <- mean(s_lenData_OJ_20)
sd_OJ_20 <- sd(s_lenData_OJ_20)

##VC - ascorbic acid
s_lenData_VC_05 <- rawData[rawData$supp == "VC" & rawData$dose == 0.5,]$len
n_VC_05 <- length(s_lenData_VC_05)
mean_VC_05 <- mean(s_lenData_VC_05)
sd_VC_05 <- sd(s_lenData_VC_05)

s_lenData_VC_10 <- rawData[rawData$supp == "VC" & rawData$dose == 1.0,]$len
n_VC_10 <- length(s_lenData_VC_10)
mean_VC_10 <- mean(s_lenData_VC_10)
sd_VC_10 <- sd(s_lenData_VC_10)

s_lenData_VC_20 <- rawData[rawData$supp == "VC" & rawData$dose == 2.0,]$len
n_VC_20 <- length(s_lenData_VC_20)
mean_VC_20 <- mean(s_lenData_VC_20)
sd_VC_20 <- sd(s_lenData_VC_20)
```


```{r confidenceIntervals}
print("OJ 05 vs. 10")
t.test(s_lenData_OJ_05, s_lenData_OJ_10, paired = FALSE, var.equal = TRUE)
print("OJ 05 vs. 20")
t.test(s_lenData_OJ_05, s_lenData_OJ_20, paired = FALSE, var.equal = TRUE)
print("OJ 10 vs. 20")
t.test(s_lenData_OJ_10, s_lenData_OJ_20, paired = FALSE, var.equal = TRUE)

print("VC 05 vs. 10")
t.test(s_lenData_VC_05, s_lenData_VC_10, paired = FALSE, var.equal = TRUE)
print("VC 05 vs. 20")
t.test(s_lenData_VC_05, s_lenData_VC_20, paired = FALSE, var.equal = TRUE)
print("VC 10 vs. 20")
t.test(s_lenData_VC_10, s_lenData_VC_20, paired = FALSE, var.equal = TRUE)

print("OJ 05 vs. VC 05")
t.test(s_lenData_OJ_05, s_lenData_VC_05, paired = FALSE, var.equal = TRUE)
print("OJ 10 vs.VC 10")
t.test(s_lenData_OJ_10, s_lenData_VC_10, paired = FALSE, var.equal = TRUE)
print("OJ 20 vs. VC 20")
t.test(s_lenData_OJ_20, s_lenData_VC_20, paired = FALSE, var.equal = TRUE)
```