---
title: "ToothGrowth Data Analysis"
author: "Pier Lorenzo Paracchini"
output: html_document
---
```{r settings, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
Sys.setlocale("LC_ALL", "C")
```
##Overview
The goal is to analyze the ToothGrowth data in the R datasets package, performing some exploratory analysis, and then use confidence intervals and/or hypothesis tests to compare tooth growth by `supp` and `dose`. 

Description of the data from the documentation _"The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid)."_

##Getting and Exploring the data
```{r dataStructure, collpase = TRUE}
require(datasets)
rawData <- ToothGrowth
str(rawData)
```
The rawdata includes `r dim(rawData)[1]` obs. of `r dim(rawData)[2]` features. Features are `len` the tooth length (`numeric`), `supp` the supplement type (`factor`) and finally `dose` the dose in milligrams (`numeric`). There are 6 possible groups by `supp` and `dose` (see the plot below). For each supplement `supp` ("OJ" and "VC"), there are three different groups based on the `dose` values - 0.5mg, 1mg, 2mg.

```{r ref.label="basicPlot", echo=FALSE}
```

And here some basic statics about the sample tooth length for each group.
```{r ref.label="prepareDataGroups", echo=FALSE}
```
```{r ref.label="basicStatistics", echo=FALSE}
```
Supp | Dose | No of Sample | Sample mean | Sample SD 
---- | ---- | ---- | ---- | ----
"OJ" | 0.5 | `r stats_OJ_05$n` | `r stats_OJ_05$mean` | `r stats_OJ_05$sd` 
"OJ" | 1.0 | `r stats_OJ_10$n` | `r stats_OJ_10$mean` | `r stats_OJ_10$sd`
"OJ" | 2.0 | `r stats_OJ_20$n` | `r stats_OJ_20$mean` | `r stats_OJ_20$sd`
--- | --- | --- | --- | ---
"VC" | 0.5 | `r stats_VC_05$n` | `r stats_VC_05$mean` | `r stats_VC_05$sd` 
"VC" | 1.0 | `r stats_VC_10$n` | `r stats_VC_10$mean` | `r stats_VC_10$sd`
"VC" | 2.0 | `r stats_VC_20$n` | `r stats_VC_20$mean` | `r stats_VC_20$sd` 

##Confidence Intervals & Hypothesis Testing
###Basic Assumptions
Having __each group has a limited number of sample (10 sample)__ then __`t distribution` and `t confidence intervals`__ are going to be used. Being the guinea pigs used for each measurement distinct from each other within a group and between the groups then the __groups can be considered independent from each other__. Finally looking at the __sample standard deviation__ for the different groups the assumption can be made to treat __diffent variances as equals__. 

###Independent group t confident intervals
```{r ref.label="tTests", echo=FALSE}
```

Group1 | Group2 | t confidence interval (g1, g2)  
---- | ---- | ----
OJ, 1.0 | OJ, 0.5 | `r t.testOJ_10_vs_05$conf` 
OJ, 2.0 | OJ, 0.5 | `r t.testOJ_20_vs_05$conf` 
OJ, 2.0 | OJ, 1.0 | `r t.testOJ_20_vs_10$conf`
--- | --- | ---
VC, 1.0 | VC, 0.5 | `r t.testVC_10_vs_05$conf` 
VC, 2.0 | VC, 0.5 | `r t.testVC_20_vs_05$conf` 
VC, 2.0 | VC, 1.0 | `r t.testVC_20_vs_10$conf`
--- | --- | ---
OJ, 0.5 | VC, 0.5 | `r t.testVC_10_vs_05$conf` 
OJ, 1.0 | VC, 1.0 | `r t.testVC_20_vs_05$conf` 
OJ, 2.0 | VC, 2.0 | `r t.testVC_20_vs_10$conf`
--- | --- | ---
OJ, all | VC, all | `r t.testOJ_vs_VC$conf`

##Appendix
__Code used to create the basic plot to visualize basic information of the rawData.__
```{r basicPlot}
require(graphics)
rawData_OJ <- rawData[rawData$supp == "OJ",]
rawData_OJ$doseF <- as.factor(rawData_OJ$dose)
rawData_VC <- rawData[rawData$supp == "VC",]
rawData_VC$doseF <- as.factor(rawData_VC$dose)

par(ps = 8, cex.lab = 0.8, mfcol = c(1,2))
plot(rawData_OJ$dose, rawData_OJ$len, col = c(rawData_OJ$doseF), pch = 20, main ="Length vs. Dose for Orange Juice Supplement", xlab = "mg", ylab ="Tooth Length", ylim=c(0, 40))
legend("bottomright",cex = 0.6, fill = c("black","red","green"), legend = c("0.5mg","1mg","2mg"))

plot(rawData_VC$dose, rawData_VC$len, col = c(rawData_VC$doseF), pch = 20, main ="Length vs. Dose for Ascorbic Acid Supplement", xlab = "mg", ylab ="Tooth Length", , ylim=c(0, 40))
legend("bottomright",cex = 0.6, fill = c("black","red","green"), legend = c("0.5mg","1mg","2mg"))
```

__Code used to prepare the data of the different groups__
```{r prepareDataGroups}
len_OJ <- rawData[rawData$supp == "OJ",]$len
len_OJ_05 <- rawData[rawData$supp == "OJ" & rawData$dose == 0.5,]$len
len_OJ_10 <- rawData[rawData$supp == "OJ" & rawData$dose == 1.0,]$len
len_OJ_20 <- rawData[rawData$supp == "OJ" & rawData$dose == 2.0,]$len

len_VC <- rawData[rawData$supp == "VC",]$len
len_VC_05 <- rawData[rawData$supp == "VC" & rawData$dose == 0.5,]$len
len_VC_10 <- rawData[rawData$supp == "VC" & rawData$dose == 1.0,]$len
len_VC_20 <- rawData[rawData$supp == "VC" & rawData$dose == 2.0,]$len
```

__Code used to calculate basic statistics for the different groups__
```{r basicStatistics}
getStatistics <- function(data){
  return (list(n = length(data), mean = mean(data), sd = round(sd(data),digits = 3)))
}

##OJ Statistics
stats_OJ_05 <- getStatistics(len_OJ_05)
stats_OJ_10 <- getStatistics(len_OJ_10)
stats_OJ_20 <- getStatistics(len_OJ_20)

##VC Statistics
stats_VC_05 <- getStatistics(len_VC_05)
stats_VC_10 <- getStatistics(len_VC_10)
stats_VC_20 <- getStatistics(len_VC_20)
```

__Code used to calculate t test between the different groups__
```{r tTests}
t.testOJ_10_vs_05 <- t.test(len_OJ_10, len_OJ_05, paired = FALSE, var.equal = TRUE)
t.testOJ_20_vs_05 <- t.test(len_OJ_20, len_OJ_05, paired = FALSE, var.equal = TRUE)
t.testOJ_20_vs_10 <- t.test(len_OJ_20, len_OJ_10, paired = FALSE, var.equal = TRUE)

t.testVC_10_vs_05 <- t.test(len_VC_10, len_VC_05, paired = FALSE, var.equal = TRUE)
t.testVC_20_vs_05 <-t.test(len_VC_20, len_VC_05, paired = FALSE, var.equal = TRUE)
t.testVC_20_vs_10 <-t.test(len_VC_20, len_VC_10, paired = FALSE, var.equal = TRUE)

t.testOJ_05_vs_VC_05 <- t.test(len_OJ_05, len_VC_05, paired = FALSE, var.equal = TRUE)
t.testOJ_10_vs_VC_10 <- t.test(len_OJ_10, len_VC_10, paired = FALSE, var.equal = TRUE)
t.testOJ_20_vs_VC_20 <- t.test(len_OJ_20, len_VC_20, paired = FALSE, var.equal = TRUE)

t.testOJ_vs_VC <- t.test(len_OJ, len_VC, paired = FALSE, var.equal = TRUE)
```

##T Be removed
```{r basicPlot1}
require(graphics)
par(ps = 8, cex.lab = 0.8)
coplot(len ~ dose | supp, data = rawData, panel = panel.smooth, xlab = "ToothGrowth data: length vs dose, given type of supplement")
```

```{r basicStatistics1}
##OJ - Orange Juice
s_lenData_OJ_05 <- rawData[rawData$supp == "OJ" & rawData$dose == 0.5,]$len
n_OJ_05 <- length(s_lenData_OJ_05)
mean_OJ_05 <- mean(s_lenData_OJ_05)
sd_OJ_05 <- sd(s_lenData_OJ_05)

s_lenData_OJ_10 <- rawData[rawData$supp == "OJ" & rawData$dose == 1.0,]$len
n_OJ_10 <- length(s_lenData_OJ_10)
mean_OJ_10 <- mean(s_lenData_OJ_10)
sd_OJ_10 <- sd(s_lenData_OJ_10)

s_lenData_OJ_20 <- rawData[rawData$supp == "OJ" & rawData$dose == 2.0,]$len
n_OJ_20 <- length(s_lenData_OJ_20)
mean_OJ_20 <- mean(s_lenData_OJ_20)
sd_OJ_20 <- sd(s_lenData_OJ_20)

##VC - ascorbic acid
s_lenData_VC_05 <- rawData[rawData$supp == "VC" & rawData$dose == 0.5,]$len
n_VC_05 <- length(s_lenData_VC_05)
mean_VC_05 <- mean(s_lenData_VC_05)
sd_VC_05 <- sd(s_lenData_VC_05)

s_lenData_VC_10 <- rawData[rawData$supp == "VC" & rawData$dose == 1.0,]$len
n_VC_10 <- length(s_lenData_VC_10)
mean_VC_10 <- mean(s_lenData_VC_10)
sd_VC_10 <- sd(s_lenData_VC_10)

s_lenData_VC_20 <- rawData[rawData$supp == "VC" & rawData$dose == 2.0,]$len
n_VC_20 <- length(s_lenData_VC_20)
mean_VC_20 <- mean(s_lenData_VC_20)
sd_VC_20 <- sd(s_lenData_VC_20)
```