---
title: "Statistical_Inference"
author: "ppar"
date: "02. nov. 2015"
---
#Week 1
##Probability Density Function

Let's imagine that the proportion of help calls that get addressed in a random day by a help line is given by the PDF below

```{r pdf}
x <- c(-0.5,0,1,1,1.5)
y <- c(0,0,2,0,0)

plot(x, y, lwd = 3, frame = TRUE, type="l")
```

The PDF valid

* values are equal or greater then 0
* total area is 1

What is the probability that 75% or fewer of call get addressed?
Such propability can be calculated as the area of the triangle highlited in the diagram below
```{r collapse = TRUE}
plot(x, y, lwd = 3, frame = TRUE, type="l")
lines(c(0.0, 0.75), c(0, 0), col = "red")
lines(c(0.75, 0.75), c(0.0, 1.5), col = "red")
lines(c(0.0, 0.75), c(0.0, 1.5), col = "red")

calc_prob <- (0.75 * 1.5) / 2
calc_prob
```
 The given PDF is a special case of a well-known probaility distibution, the **BETA distribution**. The sameprobability can be calculated using the **BETA distribution** as in the following code chunk
```{r collapse = TRUE}
pbeta(0.75,2,1)
```

Some more details about the **BETA** distribution
```{r}
x <- seq(0, 1, length = 100)
d <- dbeta(x, 2, 1) #density
p <- pbeta(x, 2, 1) #probability of being less than x
plot(x, d, type = "l")
lines(x, p, type= "l", col = "red")
```
