---
title: "Statistical_Inference"
author: "ppar"
date: "02. nov. 2015"
---
#Week 1
##Probability Density Function

Let's imagine that the proportion of help calls that get addressed in a random day by a help line is given by the PDF below

```{r pdf}
x <- c(-0.5,0,1,1,1.5)
y <- c(0,0,2,0,0)

plot(x, y, lwd = 3, frame = TRUE, type="l")
```

The PDF valid

* values are equal or greater then 0
* total area is 1

What is the probability that 75% or fewer of call get addressed?
Such propability can be calculated as the area of the triangle highlited in the diagram below
```{r collapse = TRUE}
plot(x, y, lwd = 3, frame = TRUE, type="l")
lines(c(0.0, 0.75), c(0, 0), col = "red")
lines(c(0.75, 0.75), c(0.0, 1.5), col = "red")
lines(c(0.0, 0.75), c(0.0, 1.5), col = "red")

calc_prob <- (0.75 * 1.5) / 2
calc_prob
```
 The given PDF is a special case of a well-known probaility distibution, the **BETA distribution**. The sameprobability can be calculated using the **BETA distribution** as in the following code chunk
```{r collapse = FALSE}
pbeta(0.75,2,1)

#What's the probability that 40%, 50% and 60% of the calls or fewere are addressed on the same day
#F(x), given x = 0.4, 0.5, 0.6
pbeta(c(0.4, 0.5, 0.6), 2, 1)

#And the survival function is 1 - F(x)
1 - pbeta(c(0.4, 0.5, 0.6), 2, 1)
```

Some more details about the **BETA** distribution
```{r}
x <- seq(0, 1, length = 100)
d <- dbeta(x, 2, 1) #density
p <- pbeta(x, 2, 1) #probability of being less than x
plot(x, d, type = "l")
lines(x, p, type= "l", col = "red")
```

Qualtiles/ percentiles for the **BETA** distribution can be calculated 
```{r}
qbeta(0.5, 2, 1) #0.5 quantile, 50% percentile or population median
```

###Expected Values
####Find the center of mass of the bars
```{r}
library(UsingR)
library(ggplot2)
library(reshape2)
data(galton)

longGalton <- melt(galton, measure.vars = c("child", "parent"))
g <- ggplot(longGalton, aes(x = value)) +
  geom_histogram(aes(y = ..density.., fill=variable), binwidth=1, colour="black") +
  geom_density(size=2)
g <- g + facet_grid(. ~ variable)
g
```

Using `manipulate` package ...
```{r}
library(manipulate)
myHist <- function(mu){
    g <- ggplot(galton, aes(x = child))
    g <- g + geom_histogram(fill = "salmon", 
      binwidth=1, aes(y = ..density..), colour = "black")
    g <- g + geom_density(size = 2)
    g <- g + geom_vline(xintercept = mu, size = 2)
    mse <- round(mean((galton$child - mu)^2), 3)  
    g <- g + labs(title = paste('mu = ', mu, ' MSE = ', mse))
    g
}
manipulate(myHist(mu), mu = slider(62, 74, step = 0.5))
```
#t confidence interval
##Ex1
Load the data set mtcars in the datasets R package. Calculate a 95% confidence interval to
the nearest MPG for the variable mpg.

We know just few obs. (30 obs.) and we do not have any information about the distribution of the original population. Let´s assume that the underlying data are iid Gaussian/ normal so we can assumen that $$ \frac{\bar X - \mu}{S/\sqrt{n}} $$ foloows a t distribution with n-1 degrees of freedom.


```{r ex1, collapse=TRUE}
library(datasets)
data(mtcars)
str(mtcars)
#32 obs and 11 variables

sampleMean <- mean(mtcars$mpg)
sampleSd <- sd(mtcars$mpg)
n <- length(mtcars$mpg)
#intervals 
round(sampleMean + c(-1,1) * qt(0.975,n) * (sampleSd / sqrt(n)))

#or using t.test
round(t.test(mtcars$mpg)$conf.int)

```

##Ex2
Suppose that standard deviation of 9 paired differences is $1$. What value would the average difference have to be so that the lower endpoint of a 95% students t confidence interval touches
zero?

```{r ex2}
n <- 9 #obs.
sampleMean <- (1 / sqrt(n)) * qt(0.975,n-1)
```

##Ex3
An independent group Student's T interval is used instead of a paired T interval when:

We can't pair them if the groups are independent of each other as well as independent within themselves.

A paired interval is for paired observations.

##Ex4
Consider the mtcars dataset. Construct a 95% T interval for MPG comparing 4 to 6 cylinder cars (subtracting in the order of 4 - 6) assume a constant variance.

```{r ex4}
data(mtcars)
group1 <- mtcars$mpg[mtcars$cyl == 4]
group2 <- mtcars$mpg[mtcars$cyl == 6]

round(t.test(group1,group2, paired = FALSE, var.equal = TRUE)$conf.int, 1)
```
Calculated confidence interval is
3.2 10.7
attr(,"conf.level")
[1] 0.95

Refer back to comparing MPG for 4 versus 6 cylinders. What do you conclude?
The interval is above zero, suggesting 4 is better than 6 in the terms of MPG

##Ex5
Suppose that 18 obese subjects were randomized, 9 each, to a new diet pill and a placebo. Subjects' body mass indices (BMIs) were measured at a baseline and again after having received the treatment or placebo for four weeks. The average difference from follow-up to the baseline (followup - baseline) was 3 kg/m2 for the treated group and 1 kg/m2 for the placebo group. The corresponding standard deviations of the differences was 1.5 kg/m2 for the treatment group and 1.8 kg/m2 for the placebo group. The study aims to answer whether the change in BMI over the four week period appear to differ between the treated and placebo groups.
What is the pooled variance estimate? (to 2 decimal places)

```{e ex5}
n <- 18
#group1 (g1) new diet pill, group2 (g2) placebo
n_g1 <- 9 
n_g2 <- 9
sampleMeanDiff_g1 <- 3
sampleMeanDiff_g2 <- 1
sampleSdDiff_g1 <- 1.5
sampleSdDiff_g2 <- 1.8

pve <- ((n_g1 -1) * sampleSdDiff_g1^2 + (n_g2 -1) * sampleSdDiff_g2^2)/(n_g1 + n_g2 -2)
round(pve,2)
#The sample sizes are equal, so the pooled variance is the average of the individual variances
```